---
title: "Actual and desired time spent on activities"
question_id: likert01
---

```{r}
#| include: false
#| file: ../R/_load_common.R
```

```{r}
#| echo: false
nordic_countries <- c(
  "Finland", "Norway", "Sweden", "Denmark", "Iceland", "Estonia"
)
country_groups <- list(
  Nordics = nordic_countries,
  Germany = "Germany",
  Netherlands = "Netherlands"
)
plot_countries <- countries_from_groups(country_groups)

df_allocation <- load_filtered_tf(plot_countries)

nordics_fig_height <- likert_allocation_fig_height(
  likert_allocation_n_respondents(df_allocation, nordic_countries)
)
```

```{r}
#| echo: false
#| results: asis

actual_cols <- question_columns("likert0", names(df_allocation))
desired_cols <- question_columns("likert1", names(df_allocation))

actual_question <- cols |>
  dplyr::filter(.data$New_name %in% actual_cols) |>
  dplyr::pull(.data$Question) |>
  unique() |>
  stringr::str_remove_all("\\[.*?\\]") |>
  trimws()

desired_question <- cols |>
  dplyr::filter(.data$New_name %in% desired_cols) |>
  dplyr::pull(.data$Question) |>
  unique() |>
  stringr::str_remove_all("\\[.*?\\]") |>
  trimws()

activity_labels <- purrr::map_chr(actual_cols, question_label, meta = cols)

cat("### Survey question and tasks\n\n")
if (length(actual_question) > 0L) {
  cat("- **Actual:** ", actual_question[[1]], "\n", sep = "")
}
if (length(desired_question) > 0L) {
  cat("- **Desired:** ", desired_question[[1]], "\n\n", sep = "")
}
cat("Tasks included:\n")
cat(paste0("- ", activity_labels), sep = "\n")
cat("\n\n")
```

```{r}
#| echo: false

render_allocation_summary_table <- function(
    summary_tbl,
    title,
    subtitle,
    group_col = NULL,
    group_label = NULL,
    group_order = NULL
) {
  table_df <- summary_tbl |>
    dplyr::mutate(activity = as.character(.data$activity))

  if (is.null(group_col)) {
    table_df <- table_df |>
      dplyr::select(activity, actual, desired, change)

    gt_tbl <- table_df |>
      gt::gt() |>
      gt::cols_label(
        activity = "Activity",
        actual = "Actual (%)",
        desired = "Desired (%)",
        change = "Desired - actual"
      )
  } else {
    if (is.null(group_label)) {
      group_label <- "Group"
    }
    if (is.null(group_order)) {
      group_order <- unique(as.character(table_df[[group_col]]))
    }
    table_df <- table_df |>
      dplyr::mutate(
        !!rlang::sym(group_col) := as.character(.data[[group_col]])
      ) |>
      dplyr::select(dplyr::all_of(group_col), activity, actual, desired, change)

    display_df <- table_df |>
      dplyr::select(activity, actual, desired, change)
    group_vec <- table_df[[group_col]]
    row_groups <- split(seq_len(nrow(display_df)), group_vec)

    gt_tbl <- display_df |>
      gt::gt() |>
      gt::cols_label(
        activity = "Activity",
        actual = "Actual (%)",
        desired = "Desired (%)",
        change = "Desired - actual"
      )

    for (group_name in group_order) {
      if (!group_name %in% names(row_groups)) {
        next
      }
      gt_tbl <- gt_tbl |>
        gt::tab_row_group(
          label = group_name,
          rows = row_groups[[group_name]]
        )
    }
  }

  gt_tbl |>
    gt::tab_header(
      title = title,
      subtitle = subtitle
    ) |>
    gt::fmt_number(
      columns = c(actual, desired, change),
      decimals = 1
    ) |>
    gt::tab_style(
      style = gt::cell_text(weight = "bold"),
      locations = gt::cells_column_labels()
    ) |>
    gt::tab_options(table.font.size = gt::px(14))
}

render_allocation_age_wide_table <- function(summary_tbl, title, subtitle) {
  age_levels <- c("Below 35", "35-45", "45+")
  age_tables <- purrr::set_names(
    purrr::map(age_levels, function(age_level) {
      summary_tbl |>
        dplyr::filter(.data$country == age_level) |>
        dplyr::arrange(.data$activity) |>
        dplyr::mutate(activity = as.character(.data$activity)) |>
        dplyr::select(activity, actual, desired, change)
    }),
    age_levels
  )

  n_rows <- max(purrr::map_int(age_tables, nrow))
  if (n_rows == 0L) {
    stop("No age-group allocation summaries available for the requested selection.")
  }

  pad_tbl <- function(tbl, n) {
    if (nrow(tbl) >= n) {
      return(tbl)
    }
    dplyr::bind_rows(
      tbl,
      tibble::tibble(
        activity = rep("", n - nrow(tbl)),
        actual = rep(NA_real_, n - nrow(tbl)),
        desired = rep(NA_real_, n - nrow(tbl)),
        change = rep(NA_real_, n - nrow(tbl))
      )
    )
  }

  age_tables <- purrr::map(age_tables, pad_tbl, n = n_rows)

  wide_df <- tibble::tibble(
    activity_below_35 = age_tables[["Below 35"]]$activity,
    actual_below_35 = age_tables[["Below 35"]]$actual,
    desired_below_35 = age_tables[["Below 35"]]$desired,
    change_below_35 = age_tables[["Below 35"]]$change,
    activity_35_45 = age_tables[["35-45"]]$activity,
    actual_35_45 = age_tables[["35-45"]]$actual,
    desired_35_45 = age_tables[["35-45"]]$desired,
    change_35_45 = age_tables[["35-45"]]$change,
    activity_45_plus = age_tables[["45+"]]$activity,
    actual_45_plus = age_tables[["45+"]]$actual,
    desired_45_plus = age_tables[["45+"]]$desired,
    change_45_plus = age_tables[["45+"]]$change
  )

  wide_df |>
    gt::gt() |>
    gt::tab_header(
      title = title,
      subtitle = subtitle
    ) |>
    gt::tab_spanner(
      label = "Below 35",
      columns = c(activity_below_35, actual_below_35, desired_below_35, change_below_35)
    ) |>
    gt::tab_spanner(
      label = "35-45",
      columns = c(activity_35_45, actual_35_45, desired_35_45, change_35_45)
    ) |>
    gt::tab_spanner(
      label = "45+",
      columns = c(activity_45_plus, actual_45_plus, desired_45_plus, change_45_plus)
    ) |>
    gt::cols_label(
      activity_below_35 = "Activity",
      actual_below_35 = "Actual (%)",
      desired_below_35 = "Desired (%)",
      change_below_35 = "Desired - actual",
      activity_35_45 = "Activity",
      actual_35_45 = "Actual (%)",
      desired_35_45 = "Desired (%)",
      change_35_45 = "Desired - actual",
      activity_45_plus = "Activity",
      actual_45_plus = "Actual (%)",
      desired_45_plus = "Desired (%)",
      change_45_plus = "Desired - actual"
    ) |>
    gt::cols_width(
      activity_below_35 ~ gt::px(170),
      activity_35_45 ~ gt::px(240),
      activity_45_plus ~ gt::px(170)
    ) |>
    gt::fmt_number(
      columns = c(
        actual_below_35, desired_below_35, change_below_35,
        actual_35_45, desired_35_45, change_35_45,
        actual_45_plus, desired_45_plus, change_45_plus
      ),
      decimals = 1
    ) |>
    gt::sub_missing(
      columns = c(
        actual_below_35, desired_below_35, change_below_35,
        actual_35_45, desired_35_45, change_35_45,
        actual_45_plus, desired_45_plus, change_45_plus
      ),
      missing_text = ""
    ) |>
    gt::tab_style(
      style = gt::cell_text(weight = "bold"),
      locations = gt::cells_column_labels()
    ) |>
    gt::tab_options(table.font.size = gt::px(14))
}
```

## Nordics

### Descriptive Statistics

```{r}
#| echo: false
#| results: asis

actual_cols <- question_columns("likert0", names(df_allocation))
desired_cols <- question_columns("likert1", names(df_allocation))

if (
  !has_likert_observations(df_allocation, actual_cols) ||
    !has_likert_observations(df_allocation, desired_cols)
) {
  cat("Nordics: No responses available for Nordic countries in the current filter.\n\n")
} else {
  summary_tbl <- prepare_likert_allocation_summary(
    df_allocation,
    nordic_countries,
    meta = cols,
    stat = "mean"
  )

  render_allocation_summary_table(
    summary_tbl = summary_tbl,
    title = "Time allocation (Nordics)",
    subtitle = "Mean percentages by activity"
  )
}
```

For each activity, the chart shows actual time as the baseline (black square) and the corresponding desired level (black triangle) after rescaling each respondent to 100% and computing `desired − actual`. Lines always run from the actual baseline to `actual + change`; **solid** lines indicate an increase, **dotted** lines a decrease.

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 5
#| results: asis

print_likert_allocation_summary(
  df_allocation,
  nordic_countries,
  meta = cols,
  country_label = "Nordics",
  stat = "mean",
  empty_message = "No responses available for Nordic countries in the current filter."
)
```

### Statistics by age group

```{r}
#| echo: false
#| results: asis

df_nordics_age <- df_allocation |>
  dplyr::filter(.data$socio1_0 %in% nordic_countries) |>
  assign_age_groups_two() |>
  dplyr::mutate(socio1_0 = .data$age_group)

actual_cols <- question_columns("likert0", names(df_nordics_age))
desired_cols <- question_columns("likert1", names(df_nordics_age))

if (
  !has_likert_observations(df_nordics_age, actual_cols) ||
    !has_likert_observations(df_nordics_age, desired_cols)
) {
  cat("Nordics by age group: No responses available for the current filter.\n\n")
} else {
  summary_tbl_age <- prepare_likert_allocation_summary(
    df_nordics_age,
    c("Below 35", "35-45", "45+"),
    meta = cols,
    stat = "mean",
    by_country = TRUE
  )

  render_allocation_age_wide_table(
    summary_tbl = summary_tbl_age,
    title = "Time allocation by age group (Nordics)",
    subtitle = "Mean percentages by activity within each age group"
  )
}
```

### Participant allocation

::: {.callout-note}
## Rescaled time allocations (\*)

In the charts below, a bracketed asterisk `(*)` at the end of a bar marks respondents whose reported activity shares did not sum to 100%. For those individuals, each share was rescaled proportionally (multiplied by \(100 / \text{total}\)) so the bar reaches 100% while preserving the relative split across tasks.
:::

Each row is one respondent; bar segments show time shares per activity (rescaled to 100% where needed; see note above).
Respondents are sorted by their largest single-task share in actual time (specialists at the top); the same row order is used for desired time.

```{r}
#| echo: false
#| fig-width: 16
#| fig-height: !expr nordics_fig_height
#| results: asis

print_likert_allocation_paired(
  df_allocation,
  nordic_countries,
  meta = cols,
  country_label = "Nordics",
  empty_message = "No responses available for Nordic countries in the current filter."
)
```

## Within Nordics

### Descriptive Statistics

```{r}
#| echo: false
#| results: asis

actual_cols <- question_columns("likert0", names(df_allocation))
desired_cols <- question_columns("likert1", names(df_allocation))

if (
  !has_likert_observations(df_allocation, actual_cols) ||
    !has_likert_observations(df_allocation, desired_cols)
) {
  cat("Within Nordics: No responses available for Nordic countries in the current filter.\n\n")
} else {
  summary_tbl_within <- prepare_likert_allocation_summary(
    df_allocation,
    nordic_countries,
    meta = cols,
    stat = "mean",
    by_country = TRUE
  )

  render_allocation_summary_table(
    summary_tbl = summary_tbl_within,
    title = "Time allocation within Nordic countries",
    subtitle = "Mean percentages by activity and country",
    group_col = "country",
    group_label = "Country",
    group_order = nordic_countries
  )
}
```

The same summary with one line per Nordic country within each activity (tasks on the y-axis):

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 14
#| results: asis

print_likert_allocation_summary_by_country(
  df_allocation,
  nordic_countries,
  group_label = "Nordics",
  meta = cols,
  stat = "mean",
  empty_message = "No responses available for Nordic countries in the current filter."
)
```

## Between Countries

### Descriptive Statistics

```{r}
#| echo: false
#| results: asis

df_groups <- df_allocation |>
  dplyr::mutate(
    socio1_0 = dplyr::case_when(
      .data$socio1_0 %in% nordic_countries ~ "Nordics",
      .data$socio1_0 == "Germany" ~ "Germany",
      .data$socio1_0 == "Netherlands" ~ "Netherlands",
      TRUE ~ .data$socio1_0
    )
  )

actual_cols <- question_columns("likert0", names(df_groups))
desired_cols <- question_columns("likert1", names(df_groups))

if (
  !has_likert_observations(df_groups, actual_cols) ||
    !has_likert_observations(df_groups, desired_cols)
) {
  cat("Country groups: No responses available for the selected country groups.\n\n")
} else {
  summary_tbl_between <- prepare_likert_allocation_summary(
    df_groups,
    names(country_groups),
    meta = cols,
    stat = "mean",
    by_country = TRUE
  )

  render_allocation_summary_table(
    summary_tbl = summary_tbl_between,
    title = "Time allocation by country groups",
    subtitle = "Mean percentages by activity and country group",
    group_col = "country",
    group_label = "Country group",
    group_order = names(country_groups)
  )
}
```

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 6
#| results: asis

df_groups <- df_allocation |>
  dplyr::mutate(
    socio1_0 = dplyr::case_when(
      .data$socio1_0 %in% nordic_countries ~ "Nordics",
      .data$socio1_0 == "Germany" ~ "Germany",
      .data$socio1_0 == "Netherlands" ~ "Netherlands",
      TRUE ~ .data$socio1_0
    )
  )

print_likert_allocation_summary_by_country(
  df_groups,
  names(country_groups),
  group_label = "Country groups",
  meta = cols,
  stat = "mean",
  empty_message = "No responses available for the selected country groups."
)
```
